\documentclass[12pt]{article}
\usepackage{caption}
\usepackage{float}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{mdframed}
\usepackage{tabto}
\usepackage{amsmath}
\author{Aldar Saranov}
\date{\today}
\title{Development of an automatically configurable ant colony optimization framework. State of the art.}

\newmdenv[
  backgroundcolor=gray!20,
  frametitle=Definition,
  skipabove=\topsep,
  skipbelow=\topsep,
]{definition}

\newmdenv[
  backgroundcolor=white!20,
  frametitle=Algorithm,
  skipabove=\topsep,
  skipbelow=\topsep,
]{algorithm}

\newcommand{\dd}[1]{\mathrm{d}#1}


%--------------------------------------------------------------------------------

\begin{document}

\maketitle 
\newpage

\tableofcontents
\newpage

\begin{abstract}
Some animal species show an extreme degree of social organization. Such species (e.g. ants) have pheromone production and detection body parts and therefore seize an ability to communicate between each other in indirect way. This concept has inspired the development of algorithms which are based on social behavior of population called ant colony optimization algorithms (ACO). These algorithms allow to solve NP-hard problems in a very efficient manner. Since these algorithms are considered metaheuristic the development of a ACO framework is the next step of formalizing of this area is to provide tools for resolving general optimization problems. This article gives the brief overview of the current ACO research area state, existing framework description and some tools which can be used for the framework automatic configuration.
\end{abstract}



\section{Introduction (1 page)}
Section descriptions.
Pheromones.
Constructive heuristics.
Solution components.
Problem models.

\input{section-2}
\input{section-3}
\input{section-4}




\section{Existing ACO framework (5 pages)}

ACOTSP unified framework was developed by IRIDIA group. It was developed with tendency towards the purpose and component generality. It separates the general structures of ACO metaheuristics from the problem-specific domain. All standard parameters can be specified ($\alpha, \beta, \rho, m, \epsilon, etc.$) plus one can set the specific parameters ($t_{max}, t_{min}, res_{it}$ - number of iterations since last found rb-solution, $res_{bf}$ - branching factor, $res_{dist}$ - distance between global-best and iteration-worst ants, $q_0$).

\subsection{Finding a better ACO configuration for the TSP}

\subsection{Finding a better ACO configuration for the QAP}

\section{Automatic configuration in IRACE (3 pages)}

\section{Conclusions}

\subsection{Possible improvements}

\begin{thebibliography}{1}

\end{thebibliography}

\end{document} 
